---
title: 'Coursera: Linear Regression Course Project'
author: "YZ"
date: "October 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Assignment Overview

Which car characteristics have a significant impact on miles per gallon (MPG)? In particular,

- Is an automatic or manual transmission better for MPG?

- Quantify the MPG difference between automatic and manual transmissions

Data source: mtcars

## Executive Summary

Based on our analysis, there are three car characteristics that determine MPG:

- Type of transmission: automotic is more taxing on MPG

- Weight of a car: heavier cars consume more gas per mile

- Acceleration time (as measured by time per 1/4 mile): faster cars have lower MPG

- Cars with maual transmission have around 3 MPG more compared to automatic transmission, keeping everything else fixed

- However, if we control for car's weight, the difference in MPG between manual and automatic varies significantly: the lighter the car, the bigger the difference in MPG between manual and automatic transmissions

## Data Overview

Source dataset *mtcars* has 11 variables:
```{r}
dim(mtcars)
```
Our main variable of interest (dependent) is *mpg*.

Our main variable of interest (independent) is *am* (Transmission (0 = automatic, 1 = manual)).

```{r}
table(mtcars$am)
```
We can quickly check if there is emperical difference between two types of transmission:
```{r}
aggregate(data = mtcars, mpg ~ am, mean)
```
Indeed, manual cars on average have 7 MPG more. However, this doesn't account for any other factors.
Is it possible that other car characteristics are more impactful on MPG? Let's take a look at MPG correlation to other available factors:
```{r}
cor(mtcars)[1,]
```

Based on the above there are a lot of other factors that are highly correlated with MPG

## What car characteristics are the strongest determinants of MPG?

After multiple itirations (see Appendix), we selected the model that contains three factors: Transmission, Weight, Time for 1/4 mile. 
```{r}
coefficients(lm(formula = mpg ~ am + wt + qsec, data = mtcars))
```

Based on the above, cars with manual Transmission have almost 3 MPG more than automatic.

However, the interaction between Transmission type and car Weight was also included due to its significance (full output in Appendix):

```{r}
mpg_lm_final <- lm(formula = mpg ~ am + wt + qsec + am*wt, data = mtcars)
coefficients(mpg_lm_final)
```

Based on this model, the difference in MPG between manual and automatic transmission is not constant. Instead, it is determined by the formula: 14.079 - 4.141 * Weight. This means that lighter cars have much higher difference at MPG compared to heavier cars. 

All coefficients, except Intercept, are significantly significant at 95% level:
```{r}
confint(mpg_lm_final)
```
We observe high multicollinearity (VIF > 2) among some predictors. But it is due to the fact that we have an interaction term.
```{r}
library(car)
vif(mpg_lm_final)
```
The residual diagnostics (see Appendix) didn't identify any issues:

- Residuals don't have any distinct patterns (chart 1) 

- For the most part they are normally distributed (with some deviations at higher quantiles) (char 2)

- Assumption of homoscedasticity does not seem violated (chart 3)

- There are no outliers that skew the model (chart 4)

Therefore, the final selected model provides reliable insights.

<P style="page-break-before: always">
## Appendix

### Model Selection: stepwise itirations

We used the following method to select a model:
```{r, results = 'hide'}
mpg_lm_base <- lm(data = mtcars,
                  mpg ~ am)
mpg_lm_full <- lm(data = mtcars,
                  mpg ~ .)
mpg_lm_step <- step(mpg_lm_base, scope=list(lower=mpg_lm_base, upper=mpg_lm_full), direction="both")
```

```{r}
mpg_lm_step
```

Final model:
```{r}
summary(mpg_lm_final)
```

<P style="page-break-before: always">
Residual Diagnostics

```{r}
par(mfrow = c(2,2), oma = c(0, 0, 0, 0))
plot(mpg_lm_final)
```